\textbf{Selective DNA and Protein Isolation from Marine Macrophyte Surfaces}

### 2021-03-25 submission/manuscript.tex

Successful amplification and sequencing of the V4 region of the 16S rRNA
gene proved that the isolated DNA can be used to estimate the microbial
epiphytic diversity. Taxonomic groups detected in this step can also be
often found in epiphytic communities associated with other macrophytes
(Burke et al., 2011b; Morrissey et al., 2019). A problem often
encountered in studies focusing on epiphytic communities is the presence
of large proportions of chloroplast 16S rRNA sequences in the pool of
amplified molecules, especially if the epiphytic DNA was isolated
without prior selection (Staufenberger et al., 2008). These sequences
can derive from host chloroplasts or from eukaryotic epiphytic
chloroplast DNA. Although the proportion of obtained chloroplast 16S
rRNA sequences in our samples was substantial, they derived almost
exclusively from eukaryotic epiphytes. High proportion of chloroplast
16S rRNA sequences in studies applying selective procedures that include
direct cellular lysis on host surfaces were observed before (Michelou et
al., 2013). It is possible that chloroplast-specific sequences even in
these studies originated from eukaryotic epiphytic cells and not from
host chloroplasts. Indeed, it is common during 16S rRNA profiling of
pelagic microbial communities to observe high proportions of chloroplast
sequences (Gilbert et al., 2009; Korlević et al., 2016). On of the
solutions to further reduce chloroplast 16S rRNA sequence contamination
is to use primers that minimise the amplification of these reads if the
sequencing and study design allow it (Hanshew et al., 2013). In
addition, a very low proportion of chloroplast 16S rRNA sequences in
samples originating from \emph{C. cylindracea} in comparison to \emph{C.
nodosa} could be explained by the presence of three introns in the gene
for 16S rRNA in some members of the genus \emph{Caulerpa} that could
hamper the amplification process (Lam and Lopez-Bautista, 2016).

>>>>

Successful amplification and sequencing of the V4 region of the 16S rRNA
gene proved that the isolated DNA can be used to estimate the microbial
epiphytic diversity. Taxonomic groups detected in this step can also be
often found in epiphytic communities associated with other macrophytes
(Burke et al., 2011b; Morrissey et al., 2019). A problem often
encountered in studies focusing on epiphytic communities is the presence
of large proportions of chloroplast 16S rRNA sequences in the pool of
amplified molecules, especially if the epiphytic DNA was isolated
without prior selection (Staufenberger et al., 2008). These sequences
can derive from host chloroplasts or from eukaryotic epiphytic
chloroplast DNA. Although the proportion of obtained chloroplast 16S
rRNA sequences in our samples was substantial, they derived almost
exclusively from eukaryotic epiphytes. High proportion of chloroplast
16S rRNA sequences in studies applying selective procedures that include
direct cellular lysis on host surfaces were observed before (Michelou et
al., 2013). It is possible that chloroplast-specific sequences even in
these studies originated from eukaryotic epiphytic cells and not from
host chloroplasts. Indeed, it is common during 16S rRNA profiling of
pelagic microbial communities to observe high proportions of chloroplast
sequences (Gilbert et al., 2009; Korlević et al., 2016). One of the
solutions to further reduce chloroplast 16S rRNA sequence contamination
is to use primers that minimise the amplification of these reads if the
sequencing and study design allow it (Hanshew et al., 2013). In
addition, a very low proportion of chloroplast 16S rRNA sequences in
samples originating from \emph{C. cylindracea} in comparison to \emph{C.
nodosa} could be explained by the presence of three introns in the gene
for 16S rRNA in some members of the genus \emph{Caulerpa} that could
hamper the amplification process (Lam and Lopez-Bautista, 2016).
